% SPDX-FileCopyrightText: Copyright (c) 2024-2026 Yegor Bugayenko
% SPDX-License-Identifier: MIT

\section{Flow Control}\label{sec:flow}

The \deff{seq} object is an abstraction of sequence of objects
  to be dataized sequentially.
For example, this code prints one message to the console
  and then terminates the program due to
  the inability to dataize the division by zero in the middle:

\begin{ffcode}
seq *
  io.stdout "Hello, world!"
  42.div 0
  io.stdout "Bye!"
\end{ffcode}

The objects \deff{true} and \deff{false} both have
  an attribute \deff{if} which is a branching mechanism,
  expecting two arguments: a ``positive'' object, and a ``negative'' object.
When being dataized, the object \ff{true.if} always returns
  the ``positive'' object, the object \ff{false.if} always
  returns the ``negative'' object.
This code gives a title to a random number:

\begin{ffcode}
ms.random.pseudo > r
io.stdout
  tt.sprintf *1
    "Coin toss: %s"
    if.
      r.gte 0.5
      "heads"
      "tails"
\end{ffcode}

The \deff{while} object is an iterating mechanism, expecting two arguments.
Both are abstract objects with one free attribute for current iteration index.
The first argument is a condition, the second one is the body.
When being dataized, the object \deff{while} repeatedly dataizes
  the body while the \deff{if} of the condition goes ``positive.''
In the end, it returns the body or \deff{false} if no iterations happened.
This code makes a few attempts to find a random number
  that is smaller than 0.1:

\begin{ffcode}
ms.random.pseudo > r
while
  r.gte 0.1 > [i] >>
  io.stdout > [i]
    tt.sprintf *1
      "Attempt #%d"
      i
\end{ffcode}

The \deff{switch} object is an abstraction of multi-case
  branching that encapsulates a \deff{tuple} of \((k, v)\) pairs;
  the dataization result is equal to \(v\) of the first pair where
  \(k\) is \deff{true}, or \deff{true} if no pair matches:

\begin{ffcode}
io.stdout
  switch *
    *
      password.eq "swordfish"
      "password is correct!"
    *
      password.eq ""
      "empty password is not allowed"
    *
      false
      "password is wrong"
\end{ffcode}
